---
title: "Task 1 Solution Proposal"
author: "Sander Eide Dahling"
date: "2023-06-18"
output:
  pdf_document: default
  word_document: default
  html_document: default
classoption: a4paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Task 1 Solution Proposal
First we need to import all relevant packages to the memory of R. I will make use of the Tidyverse in this task, as it is relatively simple.

```{r, message = FALSE}
# Loading relevant libraries into memory
library(tidyverse)
```

Before loading in the data, you need to make sure that the working directory of the program is the same as the location of the file. I make sure this is the case by running this code:

```{r}
# Getting the current path of the source file
current_path = dirname(rstudioapi::getActiveDocumentContext()$path)

# Then you are able to set the working directory as this file path
setwd(current_path)
```

In my example, the working directory will be set to **`r current_path`**. This might not look the same to you, especially if you are on a Mac.

With the working directory set up correctly, we can start by loading in the data. The data used for this task is a data set gathered from [Kaggle](https://www.kaggle.com/datasets/drahulsingh/top-largest-universities). If you are unable to download the data set from this link, you can find the data in the "data" folder in the [Task1](https://github.com/sander-ed/nhh-course-help/blob/main/GUIDES/Data-Science-Academy/R-Training-Tasks/Task1-Top-Unis/data/Top-Largest-Universities.csv) folder. 

```{r}
# Importing the data using the read_csv() function from readr

df <- read_csv("../data/Top-Largest-Universities.csv")

# Objective 1
df_1 <- df %>%
  filter(`Distance / In-Person` == "In-Person")

# Completed Objective 1
head(df_1)
```

```{r}
# Objective 2

# Here we make use of pipes and dplyr in order to summarise the data frame by Continent, and compute the mean of the enrollment for each continent. Lastly we use the arrange function in dplyr to sort the newly created column in descending order.

df_2 <- df_1 %>% 
  group_by(Continent) %>% 
  summarise(average_enrollment = mean(Enrollment)) %>% 
  arrange(desc(average_enrollment))

# Completed Objective 2
df_2
```
```{r}
# Objective 3

# Here we need to re-use some of the code from the previous task and make some changes to the summarise function.

df_3 <- df_1 %>% 
  group_by(Continent) %>% 
  summarise(average_enrollment = mean(Enrollment),
            number_of_institutions = n()) %>% 
  arrange(desc(average_enrollment))

# Completed Objective 3
df_3
```

```{r}
# Objective 4

# Our starting position is once again the data frame df_1, where we need to create a new summarisation function in order to calculate a percentage.

# Creating a copy of the data frame, and add a new column
df_4 <- df_1 %>% 
  mutate(PrivateDummy = ifelse(Affiliation == "Private", 1, 0)) %>% 
  group_by(Continent) %>% 
  summarise(private_percentage = mean(PrivateDummy)) %>% 
  arrange(desc(private_percentage))
  
# Completed Objective 4  
df_4
```


