---
title: "Task 1 Solution Proposal"
author: "Sander Eide Dahling"
date: "2023-06-18"
output:
  html_document: default
  word_document: default
  pdf_document: default
classoption: a4paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Task 1 Solution Proposal
First we need to import all relevant packages to the memory of R. I will make use of the Tidyverse in this task, as it is relatively simple.

```{r, message = FALSE}
# Loading relevant libraries into memory
library(tidyverse)
```

Before loading in the data, you need to make sure that the working directory of the program is the same as the location of the file. I make sure this is the case by running this code:

```{r}
# Getting the current path of the source file
current_path = dirname(rstudioapi::getActiveDocumentContext()$path)

# Then you are able to set the working directory as this file path
setwd(current_path)
```

In my example, the working directory will be set to **`r current_path`**. This might not look the same to you, especially if you are on a Mac.

With the working directory set up correctly, we can start by loading in the data. The data used for this task is a data set gathered from [Kaggle](https://www.kaggle.com/datasets/drahulsingh/top-largest-universities). If you are unable to download the data set from this link, you can find the data in the "data" folder in the Task1 folder. 


