---
title: "Task 2 Solution Proposal"
author: "Sander Eide Dahling"
date: "2023-06-18"
output:
  pdf_document: default
  word_document: default
  html_document: default
classoption: a4paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Task 2 Solution Proposal
First we need to import all relevant packages to the memory of R. I will make use of the Tidyverse in this task, as it is relatively simple. We will specifically use tidyr and dplyr

```{r, message = FALSE}
# Loading relevant libraries into memory
library(tidyverse)
```

Before loading in the data, you need to make sure that the working directory of the program is the same as the location of the file. I make sure this is the case by running this code:

```{r}
# Getting the current path of the source file
current_path = dirname(rstudioapi::getActiveDocumentContext()$path)

# Then you are able to set the working directory as this file path
#setwd(current_path)
```

In my example, the working directory will be set to **`r current_path`**. This might not look the same to you, especially if you are on a Mac.

With the working directory set up correctly, we can start by loading in the data. The data used for this task is a data set gathered from [Kaggle](https://www.kaggle.com/datasets/ethanmitten/premier-league-fixtures-202324). If you are unable to download the data set from this link, you can find the data in the "data" folder in the [Task2](https://github.com/sander-ed/nhh-course-help/blob/main/GUIDES/Data-Science-Academy/R-Training-Tasks/Task2-PremierLeague-23-24/data/FPL_Schedule2324.csv) folder. 

```{r}
# Loading in the data set

df <- read_csv("../data/FPL_Schedule2324.csv")

head(df)
```

Here we can see that the data is in the wide format. According to Objective 1, we must make the format "longer".

```{r}
# Use the tidyr function "pivot_longer" and specify which columns to convert to long format
df_long <- df %>% 
  pivot_longer(cols = "GW1":"GW38")

# Completed Objective 1
head(df_long)
```

```{r}
# Changing the names of the columns to make them more readable

df_renamed <- df_long %>% 
  rename(team = Team,
         game_week = name,
         opponent = value)

# Completed Objective 2
df_renamed
```

```{r}
# Using the filter function to only choose a specific set of observations
df_last_day <- df_renamed %>% 
  filter(game_week == "GW38",
         team %in% c("Arsenal", "Liverpool", "Man City"))

# Completed Objective 3
df_last_day
```












